---
title: "fire behavior model"
author: "Kaley Dodson, Alison Sells, Larissa Neilson"
date: "11/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### define variables and compose equation

damage function for each unit i at time t, where G is a growth function (this is what I think we would modify with housing parameter)
s_it+1 = G(s-x) # post treatment fuel stock
s_it # fuel stocks
x_it # amount of fuel removed

assuming fire acts similarly on all variables:
Di = Di(s_t - x_t) 
	= sum w_ij(s_jt-x_jt) where weight w_ij represents the contribution of unit j’s posttreatment fuel stock, (s_jt−x_jt), to fire damage on unit i. The bigger the weight assigned to unit j, the more important that unit’s fuel stock in determining fire damage on unit i.
*** this may be where we adjust impact of houses on damage in next cell ***

after a fire fuels are:
s_it+1 = G[(s-x)(1-Di)]


we want fire to effect fuel stock and houses (i.e. damage to homes?)
house after fire is some variation of hi,t+1 = hit(1−Di)




```{r}
matrix = matrix(c(0, 1, 1, 0, 0, 1, 1, 1, 0), ncol = 3) # base values 

set.seed(2021)
list <- st_sample(box, size = 9) # select 9 values

if 


# pick ten random points from inside box and put them in p
set.seed(2021)
list <- st_sample(box, size = 9) # select 9 values

# from here an if else statement to determine fire spread (pull from work code)
 
# maybe map those using plot? (pull from gis code) 
```



